---
title: "Introduction to R"
---

This is an RMarkdown file. It mixes code and text. We can compile it into an HTML that contains everything. To compile it, click on “Knit”.

We can execute a chuck of code by `Ctrl+Shift+Enter` (Windows) or `Command+Shift+Enter` (macOS).

See a list of shortcuts [here](https://bookdown.org/yihui/rmarkdown-cookbook/rstudio-shortcuts.html).

## Download R and Rstudio

https://cran.r-project.org/bin/windows/base/ (Windows)

https://cran.r-project.org/bin/macosx/ (macOS)

https://www.rstudio.com/products/rstudio/download/


## R basics

Everything is stored as a variable. It's different from SPSS in this respect. We can see all the variables in the "Environment". To assign something to a variable, use `<-`
```{r}
result <- 1+1
```
To view a variable
```{r}
result
#view(result)
```

|                    | R  | Python |
| ------------------ | -- | ------ |
| Addition           | +  | +      |
| Subtraction        | -  | -      |
| Multiplication     | *  | *      |
| Division           | /  | /      |
| Exponentiation     | ^  | **     |
| Variable assignment| <- | =      |

To check data type
```{r}
class(result)
```
|                    | R         | Python              |
| ------------------ | --------- | ------------------- |
| 200                | integer   | int                 |
| 200.15             | numeric   | float               |
| TRUE / FALSE       | logical   | bool (True / False) |
| "lit review"       | character | str                 |
| To check data type | `class()` | `type()`            |


Logical comparison:

|    <  less than
|    >  greater than
|    <= less than or equal to
|    >= greater than or equal to
|    == equal to
|    != not equal to

## Load packages
```{r}
library(haven) # read data in SPSS, SAS or Stata formats
library(dplyr) # data wrangling
library(ggplot2) # plots
```
The libraries are loaded successfully if there's no error message. Warnings are usually harmless. Pay attention to the conflicts.


