---
title: "More to consider for publication"
---

```{r}
options(scipen=16)
library(dplyr)
library(ggplot2)
library(scales)
library(survey)
library(tidyverse)
```

```{r}
d <- read.csv("SOSS81_clean.csv")
w <- svydesign(ids = ~1, data= d, weights = d$weight) 
```

### Is respondents' level of comfort associated with familarity with technologies?

In this dataset, we can use heard of AVs as an indicator/ proxy of familarity with technologies.
```{r}
#Heard of self driving vehicles before today: 1=yes, 0=no
driving16 <- data.frame(prop.table(svytable(~driving16, design = w))) %>% mutate(n=Freq*length(which(!is.na(d$driving16)))) %>% mutate(label=percent(Freq, accuracy = 0.1)) %>% mutate(pct=Freq*100) 

#check distribution
ggplot(driving16)+
  geom_bar(aes(x=driving16, y = pct), fill="grey", stat="identity")+
  theme_minimal()+
  guides(fill=FALSE)+
  theme(panel.grid.major.x = element_blank())+
  scale_x_discrete(name="", labels=c("1"="Yes", "0"="No"))+
  scale_y_continuous(name="Percentage of respondents")+
  ggtitle("Have you heard of self-driving vehicles before today?")+
  geom_text(aes(x=driving16, y=pct, label=label), vjust= -0.3)
```
```{r}
#simple regression (no control variables)
summary(model <- svyolr(as.factor(driving18) ~ as.factor(driving16), design = w))
nobs(model)
st <- coef(summary(model))
pval <- pnorm(abs(st[, "t value"]),lower.tail = FALSE)* 2
st <- cbind(st, "p value" = round(pval,4))
st
```
```{r}
#multiple regression
model <- svyolr(as.factor(driving18) ~ as.factor(gender)+age+as.factor(educ)+as.factor(emp)+income+hour+as.factor(driving16), design = w)
#model <- svyolr(as.factor(driving18) ~ as.factor(gender)+age+as.factor(emp)+income, design = w)
summary(model)
nobs(model) 
st <- coef(summary(model))
pval <- pnorm(abs(st[, "t value"]),lower.tail = FALSE)* 2
st <- cbind(st, "p value" = round(pval,4))
st
```

For publication, we need to decide on whether to use simple (i.e. bivariate, one independent variable) or multiple  regression (i.e. many independent variables). Either is fine but need to be consistent. Simple regressions are more common for exploratory analysis. Multiple regressions allow us to see the unique effect that each independent variable has on the outcome variable, even if the independent variables are correlated. (We need to think about what's in the literature, the research questions, and the hypothesis.)

Note on laborforce(i.e. emp): "In the labor force" include unemployed, looking for work

>This is the SPSS Syntax:
Compute laborforce=-9.
If (CD15 lt 7 or cd15=11)laborforce=1.
If (cd15 ge 7 and cd15 lt 11)laborforce=2.
Missing values laborforce (-9).
Value labels laborforce 1 'In the labor force'    2 'Not in labor force'.
Variable labels laborforce 'Is respondent in the labor force or not'.

>Which translates to:
laborforce=1 In the labor force === CD15=1  WORK FULL TIME
laborforce=1 In the labor force === CD15=2  WORK PART TIME
laborforce=1 In the labor force === CD15=3  WORK AND GO TO SCHOOL
laborforce=1 In the labor force === CD15=4  THE ARMED FORCES
laborforce=1 In the labor force === CD15=5  HAVE A JOB, BUT NOT AT WORK LAST WEEK (ON VAC, SICK LEAVE, ETC)
**laborforce=1 In the labor force === CD15=6  UNEMPLOYED, LAID OFF, LOOKING FOR WORK**
laborforce=2 Not in labor force === CD15=7  RETIRED
laborforce=2 Not in labor force === CD15=8  SCHOOL FULL TIME
laborforce=2 Not in labor force === CD15=9  HOMEMAKER
laborforce=2 Not in labor force === CD15=10  DISABLED
laborforce=1 In the labor force === CD15=11  SEMI-RETIRED, RETIRED AND WORKING PART-TIME
laborforce=-9 (Missing) === CD15=95  MISC/OTHER
laborforce=-9 (Missing) === CD15=99  REFUSED


### One plot of response distribution

```{r}
#combine into one dataframe
driving18 <- data.frame(prop.table(svytable(~driving18, design = w))) %>% mutate(n=Freq*length(which(!is.na(d$driving18)))) %>% mutate(label=percent(Freq, accuracy = 0.1)) %>% mutate(pct=Freq*100)%>% add_column(tech="Warning about slower moving vehicles") %>% mutate(comfort =driving18)
driving19 <- data.frame(prop.table(svytable(~driving19, design = w))) %>% mutate(n=Freq*length(which(!is.na(d$driving19)))) %>% mutate(label=percent(Freq, accuracy = 0.1)) %>% mutate(pct=Freq*100)%>% add_column(tech="Braking automatically") %>% mutate(comfort=driving19)
driving20 <- data.frame(prop.table(svytable(~driving20, design = w))) %>% mutate(n=Freq*length(which(!is.na(d$driving20)))) %>% mutate(label=percent(Freq, accuracy = 0.1)) %>% mutate(pct=Freq*100)%>% add_column(tech="Adaptive cruise control") %>% mutate(comfort=driving20)
driving21 <- data.frame(prop.table(svytable(~driving21, design = w))) %>% mutate(n=Freq*length(which(!is.na(d$driving21)))) %>% mutate(label=percent(Freq, accuracy = 0.1)) %>% mutate(pct=Freq*100)%>% add_column(tech="Warning about distracted driving") %>% mutate(comfort=driving21)

alltech <- bind_rows(driving18, driving19, driving20, driving21)
```
```{r}
mycolor_grey = c("grey45",  "grey60", "grey75", "grey90")
ggplot(alltech)+
  geom_bar(aes(x=factor(tech, levels = c("Braking automatically", "Adaptive cruise control","Warning about distracted driving","Warning about slower moving vehicles")),y=pct, fill=factor(comfort, levels = c(4,3,2,1))),stat="identity", position = "stack")+
  theme_minimal()+
  theme(panel.grid.major.x = element_blank(), legend.position = c(0.3, -0.2), legend.direction ="horizontal", plot.margin= grid::unit(c(0.1,0.1,2,0.1), "lines"))+
  scale_fill_manual(name="", values = mycolor_grey,labels=c("1"="Very uncomfortable", "2"="Somewhat uncomfortable", "3"="Somewhat comfortable", "4"="Very comfortable"), guide = guide_legend(reverse = TRUE))+
  scale_x_discrete(name="")+
  scale_y_continuous(name="Percentage of respondents")+
  geom_text(aes(x=factor(tech),y = pct, label= label),size = 2.5, position = position_stack(vjust = 0.5))+
  ggtitle("")+
  coord_flip()
```

```{r}
#save the plot
#ggsave("driving18_21.png", width = 8, height = 4)
```
```{r}
#check the number of respondents
colSums(!is.na(d))
```
It is usually good practice to include the number of respondents as a note under a plot.

Notice that driving22 and driving23 have poor response rate.

```{r}
d_nona <- d %>% select(caseid, weight, driving18, driving19, driving20, driving21, gender, age, agec, emp) %>% na.omit()
```


### Comparison between different techs

Let's consider this question: does respondents' level of comfort regarding driver assistance technologies differ by different types of technologies? 

Judging by the bar plot above, the answer is yes. We can quantify the difference by a model as before. Not like previous regression, here the variable of interest (level of comfort) is spread out across columns (driving18-21). So we need to reformat the data such that "level of comfort" are gathered together as one single variable (I call it "comfort").  
```{r}
# reshape wide format to long format
d_gather <- d %>% pivot_longer(c(driving18, driving19, driving20, driving21,), names_to = "tech", values_to = "comfort")
w_gather <- svydesign(ids = ~1, data= d_gather, weights = d_gather$weight) # the new survey design object
```

```{r}
# pick a tech as base
summary(model <- svyolr(as.factor(comfort) ~ relevel(as.factor(tech), ref="driving19"), design = w_gather))
st <- coef(summary(model))
pval <- pnorm(abs(st[, "t value"]),lower.tail = FALSE)* 2
st <- cbind(st, "p value" = round(pval,4))
st
```

```{r}
svyby(~as.numeric(comfort), ~as.factor(tech), w_gather, svymean, na.rm=T)
```
```{r}
# check the oddds ratio
round(exp(coef(model)), 2)
```
Compared to automatic braking (driving19), respondents are stasistically significantly more likely to be comfortable with warning about a slower moving vehicle (driving18) - the odds is 2.41 times of being comfort with automatic breaking.

If you'd like to see how to report model estimates or an example of write-up, [here](https://www.sciencedirect.com/science/article/pii/S0965856420305929) is a paper we wrote also using the SOSS data.
